\chapter*{Acknowledgements}
\addcontentsline{toc}{chapter}{Acknowledgements}
At the end of this three-year journey there are so many people that had a role in making it better and more worthwhile that I think I'm going to struggle to list them all, but I'll try my best. 
Starting from the most relevant for this work, I want to thank my supervisor, Francesco Giacomini not only for the guidance while writing this thesis, but for allowing me to have this great experience in the first place and helping me navigate towards my master's. Next up, the group I've been working with for the past few months, Patatrack, has been incredibly great in introducing me to high energy physics reconstruction. A particular thank you to Felice Pantaleo and Andrea Bocci for their invaluable guidance and desire to share their knowledge with me. As for the others, I want to mention Wahid and Tony for always providing helpful suggestions from positions I can only look up to.  I also want to thank Aurora, Nikos and JJ for being the best students-colleagues I could have wished for in this amazing experience at CERN. 
Moving closer to home, I feel like I need to start with my family. There are probably a thousand reasons to thank all of them for supporting during the last 22 years. However, I can think of three reasons in particular to thank my father and it has a lot to do with old stories and bedtime. As for my mother, she has simply always been there, through thick and thin and, no matter how hard I tried to stay miserable, has always managed to get to me and cheer me up. As per my brother, he's always been a point of reference, someone to look up to and one of my biggest motivations to move forward even when I didn't feel like I could do it. Moving a bit further away, I cannot forget my uncle (yes, yellow and red are still great colours, even in Switzerland), my aunt (and her amazing late dog) and my grandmother who can never get enough of telling me how much she believes in me and what I do. 
There are some more people who have been extremely important in the path that led me here and I want to take some time to thank them as well. Starting from the physics-related one, this thesis would probably have never been made if it weren't for professor Accorsi and his completely unique way of showing me the incredibly interesting world of physics which is hiding right behind a curtain made of math. As per the other one,  I can positively say that nothing I've written until now would have been the same if it weren't for the extremely inspiring work of professor Guastamacchia who instilled in me the love for the English language. 
It would be impossible for me not to thank two of the people who know me best, so thank you Manuel and Nello (see, I've used the short version, just like in my contact list) for all of the memories and the Fridays and, well, everything else, you probably know it better than me at this point.
Finally, there are at least 23 reasons for which I am deeply grateful to you, Francesco, but I'll keep things simple just for this time. Thank you for always being there, especially when I'm not at my best, to show me how to take life one small step at a time and for always seeing the best part of me. 